---
title: "Working With PCL Data"
author: "Jeff Atkins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{PCL Vignette}
  \usepackage[utf8]{inputenc}


---
```{r, echo = FALSE, message = FALSE}
library(plyr)
library(dplyr)
library(forestr)
```


This vignette is aimed at those working with portable canopy LiDAR (PCL) data. Raw PCL data is most typically stored in comma separated values (.csv) files that contain a series of ASCII strings of return distances and return intensities. These return distances are interspersed with data markers that correspond to known distance measures on the ground as described in Hardiman et al. (2013). The spacing between these markers, typically at 10 m, is used as an input into the forestr processing algorithm (Atkins et al. in prep). 

This vignette walks through how to work with  both single PCL files and multiple PCL files into forestr in order to calculate canopy structural metrics.

# Single Transects or Files

## Importing Data

To import a single file use the `process_pcl` command. As input, `process_pcl` takes the name and directory of the file, the height of the laser from the ground, and the marker spacing used along the transect (Atkins et al. in prep). In the example below, 1 is entered as the laser height and 10 is used as the marker spacing. Both of these measurements are in meters. The laser height adjustement is used to account for the fact that the laser is user-mounted. The laser height is added to each return distance so that each return distance represents a measurement from the ground surface, rather than the laser. The marker spacing is used in the algorithm to section out the data based on the data markers as outlined in Atkins et al. (in prep).

```{r, pcl}
forestr::process_pcl(osbs, user_height = 1, marker.spacing = 10, max.vai = 8)
```

The `process_pcl` file will print multiple CSC metrics to the screen, including canopy rugosity, rumple, porosity, etc., but will also create an output directory named output within the working directory where it will store four things:

1.  The Summary Matrix - a .csv file of columnar values of mean leaf heigh, maximum height, vegetation area index (VAI), etc.
2.  The Hit Matrix - a .csv file containing VAI by rows where each row correpsonds to the x, z position of VAI in the canopy.
3.  The Output File - this is a .csv file that contains all CSC metrics in output form.
4.  The Hit Grid - this is a PNG image file of the vertical hit grid as explained in Hardiman et al. (2013) and is a graphic representation of the vertical distribution of leaf density throughout the canopy. This can be recreated and modified using the source code with the Hit Matrix file as the input.


